---
title: "Coral Disease Research Network"
author: "LMM", "AA", "AV", "AC"
date: "25 de marzo de 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---
#1. Preparations

```{r packages & scripts}
library(igraph)
library(disparityfilter)
library(linkcomm)
library(intergraph)
#library(GGally)
library(ggpubr)
library(ggrepel)
library(cowplot)
library(viridis)
library(ggnetwork)
library(vegan)
library(dplyr)
library(strigr)
library(stringi)
source("R/cmbDFrame.R")
source("R/backbone2.R")
#source("CVsplit.R")
source("R/strenght_distribution.R")
source("R/getMaxden.R")
source("R/plotLinkCommMembers2.R")
source("R/plotLinkCommGraph2.R")
#rcomm_exx
source(url("https://raw.githubusercontent.com/alspeed09/OverlapCommAssortativity/master/R/rcomm_exx.R?token=ANC4vCmHL9P-K8CpslgXX03hOzh_fT1Tks5bXyJtwA%3D%3D"))
#xy_subsets
source(url('https://raw.githubusercontent.com/alspeed09/OverlapCommAssortativity/master/R/xy_subsets.R?token=ANC4vOheNymI7sU3IkjghOpvMAXk_vHgks5bXyRUwA%3D%3D'))
#new_exy_calculation.R
source(url("https://raw.githubusercontent.com/alspeed09/OverlapCommAssortativity/master/R/new_exy_calculation.R?token=ANC4vMi0Xiitu_JIHBrnTogYxU6zlRWOks5bXynrwA%3D%3D"))
#load permutational community assortativity
source(url('https://raw.githubusercontent.com/alspeed09/OverlapCommAssortativity/master/R/Permutational_community_assortativity.R?token=ANC4vCCSGbAfbmz5FjtwlSppKXaLBaqRks5bXyorwA%3D%3D'))
# rcomm_exx
source(url('https://raw.githubusercontent.com/alspeed09/OverlapCommAssortativity/master/R/rcomm_exx.R?token=ANC4vBY60iLlFOWDk9udRWSYcNtuclz4ks5bW4AkwA%3D%3D'))
#adj_to_edgelist
source(url('https://raw.githubusercontent.com/alspeed09/OverlapCommAssortativity/master/R/adj_to_edgelist.R?token=ANC4vFm_KUe0WpVjILXOtMMFAj2gPfpcks5bYdUdwA%3D%3D'))
```

```{r Data}
rawnet <- read.csv("data/master_codirnet.csv", sep = ",",
                   na.strings = "", header = T, stringsAsFactors = F)
coral.att <- read.csv("data/glossary_codirnet.csv") #this also provides the nodes attributes
```

```{r underscore to spaces}
rawnet <- data.frame(lapply(rawnet, function(x) {
  gsub("_", " ", x)
  }
  ), stringsAsFactors = F)

coral.att <- data.frame(lapply(coral.att, function(x) {
  gsub("_", " ", x)
  }
  ), stringsAsFactors = F)
```

coral.merge and coral.edges take the raw data file and transform it into a list of edges

```{r Edge list}
coral.merge <- cmbDFrame(rawnet[,-(1:4)])
coral.edges <- as.data.frame(coral.merge$mrgCCombs)
```


#2. Network construction

Now, we define the graph and give a weight to its edges.

```{r Main network}
coral.graph <- graph.data.frame(coral.edges, directed = F)

E(coral.graph)$weight <- rep(1, ecount(coral.graph))

coral.graph <- simplify(coral.graph,
                     edge.attr.comb = list(name="ignore", weight="sum"))

is.simple(coral.graph)

print(coral.graph)
```

Before further proceeding, we can perform the following data quality verifications

```{r verification}
#network nodes absent from the glossary
names(V(coral.graph))[!(tolower(names(V(coral.graph))) %in% tolower(as.character(coral.att$node)))]

#glossary nodes absent from the network
tolower(as.character(coral.att$node))[!(tolower(as.character(coral.att$node)) %in% tolower(names(V(coral.graph))))]

#do the network and the glossary have the same number of nodes?
length(unique(tolower(as.character(coral.att$node)))) ==
length(unique(tolower(names(V(coral.graph)))))

## are there duplicated items because of uppercase characters?
tolower(names(V(coral.graph)))[duplicated(tolower(names(V(coral.graph))))] 

```

Now we add some attributes to the nodes

```{r addition of attributes}
glossary2 <- coral.att

glossary2$node <- names(V(coral.graph))

coral.att$node <- as.character(tolower(coral.att$node))

for (i in 1:nrow(coral.att)) {
  glossary2$attribute[i] <- coral.att$attribute[glossary2$node[i] == tolower(coral.att$node)] 
}

V(coral.graph)$type <- glossary2$attribute
```

We perform our analysis over a reduced form of the network that retains the overall properties of the original graph *sensu* [Ahn et al. (2011)](https://www.nature.com/articles/srep00196)

```{r Backbone network}
coral.backbone <- graph.data.frame(backbone2(coral.graph), 
                                   directed = F)

coral.backbone <- decompose.graph(coral.backbone)[[1]]

is_simple(coral.backbone)

coral.backbone <- simplify(coral.backbone, remove.multiple = T)
```


Veamos las propiedades de la red completa

#3. Network properties

Let's examine some basic properties of our basic network

```{r basic network properties}
V(coral.graph) #number of nodes

E(coral.graph) #number of edges

graph.density(coral.graph) #network density

count_components(coral.graph) #connected components

mean(igraph::degree(coral.graph)) #mean degree

mean(strength(coral.graph)) #mean weighted degree
```

And we can compare with the values of the backbone network

```{r backbone network properties}
V(coral.backbone) #number of nodes

E(coral.backbone) #number of edges

graph.density(coral.backbone) #network density

count_components(coral.backbone) #connected components

mean(degree(coral.backbone)) #mean degree

mean(strength(coral.backbone)) #mean weighted degree
```

We can check the weighted degree distributions of each type of node

```{r weighted degree distributions}
sd_att <- data.frame(a = glossary2$attribute , 
                     b = names(strength(coral.graph)), 
                     c =  strength(coral.graph),
                     row.names = seq(1,nrow(glossary2)
                                     )
                     )
```

And visualize each distribution

```{r weighted degree plot}
plot_all <- ggplot(sd_att, aes(sd_att$c)) + 
  geom_histogram(binwidth = 40) + xlab("Vertex strength") +
  ylab("Frequency")
```


```{r diseases weighted degree}
sd_disease <- sd_att[sd_att$a=="disease",]

plot_disease <- ggplot(sd_disease, aes(sd_disease$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

```{r ecoregions weighted degree}
sd_ecoregion <- sd_att[sd_att$a=="ecoregion",]

plot_ecoregion <- ggplot(sd_ecoregion, aes(sd_ecoregion$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

```{r genera weighted degree}
sd_genera <- sd_att[sd_att$a=="genera",]

plot_genera <- ggplot(sd_genera, aes(sd_genera$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

```{r topic weighted degree}
sd_topic <- sd_att[sd_att$a=="topic",]

plot_topic <- ggplot(sd_topic, aes(sd_topic$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

We can compose the previous distributions in a single plot

```{r fig01-composed plot}
fig01top <- plot_grid(plot_all, plot_disease, labels = c("A", "B"), ncol = 2)

fig01mid <- plot_grid(plot_ecoregion, plot_genera, labels = c("C", "D"), ncol = 2)

fig01low <- plot_grid(plot_topic, labels = c("E"), ncol = 1)

dir.create("analysis/figures/")

svg("analysis/figures/fig01.svg", width = 8)

plot_grid(fig01top, fig01mid, fig01low,  nrow = 3, 
          align = "hv", rel_widths = c(1, 1, 0.1))

dev.off()
```


#4. Top topics

Given the strength distribution obtained, we can explore who are these popular topics

```{r topic categories}
ecoregions <- tolower(coral.att$node[coral.att$attribute == "ecoregion"])

diseases <- tolower(coral.att$node[coral.att$attribute == "disease"])

genus <- tolower(coral.att$node[coral.att$attribute == "genera"])

topics <- tolower(coral.att$node[coral.att$attribute == "topic"])
```

```{r topic frecuencies}
ls_ecoregions <- list()
ls_diseases   <- list()
ls_genus      <- list()
ls_topic      <- list()

for(i in 1:nrow(rawnet)) {
  ls_ecoregions[[i]] <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% ecoregions)]
  
  ls_diseases[[i]] <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% diseases)]
  
  ls_genus[[i]] <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% genus)]
  
  ls_topic[[i]] <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% topics)]
}

#number of papers by ecoregions
ecor_freq <- table(as.vector(t(do.call(cbind, ls_ecoregions)[1, ])))

#number of papers by diseases
enfe_freq <- table(as.vector(t(do.call(cbind, ls_diseases)[1, ])))

#number of papers by genera
gene_freq <- table(as.vector(t(do.call(cbind, ls_genus)[1, ])))

#number of papers by topic
topic_freq <- table(as.vector(t(do.call(cbind, ls_topic)[1, ])))
```

```{r top studied elements}
head(sort(ecor_freq*100/nrow(rawnet), decreasing = T), 5)

head(sort(enfe_freq*100/nrow(rawnet), decreasing = T), 6)[2:6]

head(sort(gene_freq*100/nrow(rawnet), decreasing = T), 6)[2:6]

head(sort(topic_freq*100/nrow(rawnet), decreasing = T), 5)
```


#5. Communities detection: edges communities

We are using the link communities approach in order to obtain overlapping communities

```{r link communities}
raw.link.com <- as.data.frame(cbind(get.edgelist(coral.backbone), 
                      E(coral.backbone)$weight))

colnames(raw.link.com) <- c("source","target","weight")

raw.link.com$weight <- as.numeric(as.character(raw.link.com$weight))

lc <- getLinkCommunities(raw.link.com)
```

If we call the *lc* object, we can see a summary of the output. Under this apporach it is possible to obtain nested communities. We could merge this cases redefining the threshold for the communities where the nested communities become part of the communities they are contained into.

```{r eliminate nested communities}
new_cut <- getMaxden(lc) #in this case, we have no nested comms
#lc <- (newLinkCommsAt(lc, cutat = new_cut[1]))
```

#6. Communities robustness

The robustness of these communities can be assessed with our proposed modification of the metric proposed by Shizuka et al. (2016). This modification takes into account the overlapping of the communities. Community structures departing from randomess will get a rcomx value close to 

```{r rcomx calculation}
Pij_comm <- calc_Pmatrix(coral.backbone, 100)

Exx <- rcomm_exx(Pij_comm)

Exy <- rcom_exy()

roc <- (Exx - Exy)/(1 - Exy)
#We got a roc value of 0.23
```


#7. Plots and figures

##7.1. Community simmilarity:


```{r simmilarity data}
comm.matrix <- getCommunityMatrix(lc, nodes = head(names(lc$numclusters), lc$numbers[2]))

comm.matrix <- t(comm.matrix)

comm.jacc <- vegdist(comm.matrix, method = "jaccard", binary = T)

comm.jacc <- (1-comm.jacc)

comm.jacc <- as.matrix(comm.jacc)

comm.graph <- graph_from_adjacency_matrix(comm.jacc, mode = "lower",
                                          diag = F, weighted = T, 
                                          add.rownames = T)

E(comm.graph)$weight2 <- E(comm.graph)$weight*10

#

comm.sizes <- data.frame(cbind(names(lc$clustsizes),lc$clustsizes), row.names = 1:length(lc$clustsizes))

colnames(comm.sizes) <- c('comm', 'size')

comm.sizes$comm <- as.numeric(as.character(comm.sizes$comm))

comm.sizes$size <- as.numeric(as.character(comm.sizes$size))

comm.sizes <- sort(comm.sizes)

V(comm.graph)$name <- c(paste("C", 1:lc$numbers[3],sep = ''))
V(comm.graph)$sizes <- comm.sizes$size
```


```{r fig02A-comm similarity graph}

comm_graph <- ggplot(data = ggnetwork(comm.graph, layout = "circle"),
       aes(x, y, xend = xend, yend = yend)) +
  geom_edges(aes(color = weight2, size = weight2*10)) +
  geom_nodes(aes(size = sizes**1.5), color = "black", fill = "black") +
  scale_color_gradient(low = "blue", high = "black") +
  geom_nodetext_repel(aes(label = vertex.names),
                      nudge_x = c(rep(0.05,8),rep(-0.05,7))) +
  scale_size_continuous(name = "Community\nsize",
                        breaks = c(4, 9, 17, 24)**1.5,
                        labels = c("4", "9", "17", "24"),
                        guide = guide_legend(override.aes = list(
                          linetype = c(rep(0, 4)))
                        )) +
  scale_color_continuous(name = "Shared nodes",
                         breaks = c(3/10, 14/10, 28/10, 46/10),
                         labels = c("3%", "14%", "28%", "46%"),
                         trans = "reverse") + theme_blank()
plot(comm_graph)

```


```{r fig02B}
plot_hubs <- data.frame(x1 = names(head(getCommunityCentrality(lc), 10)),
                          x2 = round(head(getCommunityCentrality(lc), 10), 2)) %>% 
  `row.names<-`(seq(1:10)) %>% 
  ggtexttable(rows = NULL, cols = c('Nodes', 'Community\ncentrality'),
              theme = ttheme('mBlue', base_size = 9))
  
plot_hubs
```


```{r fig02-composed}
svg(filename = "analysis/figures/fig02.svg", width = 8.5, height = 4)

plot_grid(comm_graph, plot_hubs, ncol = 2, rel_widths = c(2, 1))

dev.off()
```

##7.2. Membership matrices

The following is a global membership matrix. We'll use this as a supplementary figure.

```{r community membership plot}
svg("analysis/figures/supp_comm_membership.svg")
#win.metafile(filename="cairo", antialias="default")
comm.palette <- viridis_pal(option = "viridis")(lc$numbers[3])

plotLinkCommMembers2(lc, nodes = head(names(lc$numclusters), 90),
                    pal = comm.palette,
                    total = T, fontsize = 6, sigmafont = 4, comfont = 5,
                    sumfont = 4,
                    maxclusters = lc$numbers[3], title = F, nspace = 5)
dev.off()
```


We can plot now the membership matrices depending of the desired communities properties, e.g. we can study the communities according on their size:

Communities 3, 6, 8, 9, and, 10 are the smallest clusters

```{r Node abbreviations}
namesnode <- coral.att$node 

namesnode2 <- NULL

for(i in 1:length(namesnode))  {
  if(stri_length(namesnode[i]) < 15 | length(strsplit(namesnode[i], " ")[[1]]) == 1)  {
    namesnode2[i] <- stri_trans_totitle(namesnode[i])
  } else {
    namesnode2[i] <- strsplit(namesnode[i], " ") %>% 
    sapply(function(x){paste(stri_trans_totitle(paste(substring(x[1], 1, 15))), 
                           toupper(paste(substring(x[2:length(x)], 1, 1), collapse = "")), 
                           collapse = "")})
  }
}
```


```{r plot attributes}
coral.commconn <- data.frame(node = names(getCommunityCentrality(lc)), 
                             commconn = getCommunityCentrality(lc),
                             row.names = seq(1:lc$numbers[2]))

coral.att <- left_join(coral.att, coral.commconn, by = 'node')
coral.att$abb <- namesnode2

rm(coral.commconn)

short_pallete <- viridis_pal(option = "viridis")(7) #we added an extra color to avoid confusing commcon points
```


```{r small communities}
members_1 <- lc$nodeclusters %>%
  left_join(coral.att, by = 'node') %>%
  filter(cluster %in% c(3, 6, 8, 9, 10)) %>% 
  mutate(attribute = factor(attribute, levels = unique(attribute)),
         abb = factor(abb, levels = sort(unique(abb), decreasing = T)))


ggplot(data = members_1, 
       aes(x = cluster, y = abb, fill = cluster)) + 
  geom_tile(color = "white") +
  scale_fill_manual(values = short_pallete[2:7]) +
  facet_grid(attribute ~., scales = 'free_y', space = "free_y") +
  theme(legend.position = "none",
        strip.background = element_rect(color = "black", fill = "white"),
        strip.text = element_text(margin = margin(.2, .2, .2, .2, "cm"))) +
        geom_point(aes(x = cluster, y = abb, size = commconn), 
                   color= "white", alpha = 1/2) +
        labs(y = "Nodes", x = "Cluster")
```

```{r medium-sized communities}
members_2 <- lc$nodeclusters %>%
  left_join(coral.att, by = 'node') %>%
  filter(cluster %in% c(1, 7, 11, 13, 15)) %>% 
  mutate(attribute = factor(attribute, levels = unique(attribute)),
         abb = factor(abb, levels = sort(unique(abb), decreasing = T)))


ggplot(data = members_2, 
       aes(x = cluster, y = abb, fill = cluster)) + 
  geom_tile(color = "white") +
  scale_fill_manual(values = short_pallete[2:7]) +
  facet_grid(attribute~., scales = "free_y", space = "free_y") +
  theme(legend.position = "none",
        strip.background = element_rect(color = "black", fill = "white"),
        strip.text = element_text(margin = margin(.2, .2, .2, .2, "cm"))) +
        geom_point(aes(x = cluster, y = abb, size = commconn), 
                   color= "white", alpha = 1/2) +
        labs(y = "Nodes", x = "Cluster")
```

```{r large communities}
members_3 <- lc$nodeclusters %>%
  left_join(coral.att, by = 'node') %>%
  filter(cluster %in% c(2, 4, 5, 12, 14)) %>% 
  mutate(attribute = factor(attribute, levels = unique(attribute)),
         abb = factor(abb, levels = sort(unique(abb), decreasing = T)))


ggplot(data = members_3, 
       aes(x = cluster, y = abb, fill = cluster)) + 
  geom_tile(color = "white") +
  scale_fill_manual(values = short_pallete[2:7]) +
  facet_grid(attribute~., scales = "free_y", space = "free_y") +
  theme(legend.position = "none",
        strip.background = element_rect(color = "black", fill = "white"),
        strip.text = element_text(margin = margin(.2, .2, .2, .2, "cm"))) +
        geom_point(aes(x = cluster, y = abb, size = commconn), 
                   color= "white", alpha = 1/2) +
        labs(y = "Nodes", x = "Cluster")
```

#8. External plots

We can use certain functions to export our networks as gephi files

```{r gephi-backbone}
#sink("coral_backbone.gexf")
#igraph.to.gexf(coral.backbone)
#sink()
```


```{r gephi-main}
#sink("coral_main.gexf")
#igraph.to.gexf(coral.graph)
#sink()
```
