---
title: "cCoral Disease Research Network"
author: "LMM", "AS", "AV", "JC", "AC"
date: "25 de marzo de 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r packages & scripts}
library(igraph)
library(disparityfilter)
library(cowplot)
library(linkcomm)
library(intergraph)
library(GGally)
library(ggrepel)
library(viridis)
library(ggnetwork)
library(vegan)
source("R/cmbDFrame.R")
source("R/backbone2.R")
#source("CVsplit.R")
source("R/strenght_distribution.R")
source("R/non_nested_height2.R")
source("R/plotLinkCommMembers2.R")
source("R/plotLinkCommGraph2.R")
#devtools::install_github("thomasp85/patchwork", force=T)
#library(patchwork) #this replaces cowplot
```

```{r Data}
rawnet <- read.csv("data/master_coral_dis_net.csv", sep = ",", na.strings = "",
                   header = T, stringsAsFactors = F)
coral.att <- read.csv("data/glossary.csv") #this also provides the nodes attributes
```

```{r, underscore to spaces}
rawnet <- data.frame(lapply(rawnet, function(x) {
  gsub("_", " ", x)
  }
  ), stringsAsFactors = F)

coral.att <- data.frame(lapply(coral.att, function(x) {
  gsub("_", " ", x)
  }
  ), stringsAsFactors = F)
```

coral.merge and coral.edges take the raw data file and transform it into a list of edges

```{r Edge list}
coral.merge <- cmbDFrame(rawnet[,-(1:4)])
coral.edges <- as.data.frame(coral.merge$mrgCCombs)
```

Now, we define the graph and give a weight to the edges.

```{r Main network}
coral.graph <- graph.data.frame(coral.edges, directed = F)
E(coral.graph)$weight <- rep(1, ecount(coral.graph))
coral.graph <- simplify(coral.graph,
                     edge.attr.comb = list(name="ignore", weight="sum"))
is.simple(coral.graph)
print(coral.graph)
```

Before proceeding further, we can perform the following data quality verifications

```{r verification}
#network nodes absent from the glossary
names(V(coral.graph))[!(tolower(names(V(coral.graph))) %in% tolower(as.character(coral.att$node)))]

#glossary nodes absent from the network
tolower(as.character(coral.att$node))[!(tolower(as.character(coral.att$node)) %in% tolower(names(V(coral.graph))))]

#do the netowrk and the glossary have the same number of nodes
length(unique(tolower(as.character(coral.att$node)))) ==
length(unique(tolower(names(V(coral.graph)))))

## are there duplicated items because uppercases?
tolower(names(V(coral.graph)))[duplicated(tolower(names(V(coral.graph))))] 

```

Now we add some attributes to the nodes

```{r addition of attributes}
glossary2 <- coral.att
glossary2$node <- names(V(coral.graph))
coral.att$node <- as.character(coral.att$node)
for (i in 1:nrow(coral.att)) {
  glossary2$attribute[i] <- coral.att$attribute[glossary2$node[i] == tolower(coral.att$node)] 
}
V(coral.graph)$type <- glossary2$attribute
```

We perform our analysis over a reduced form of the network that retains the overall properties of the original graph *sensu* [Ahn et al. (2011)](https://www.nature.com/articles/srep00196)

```{r Backbone network}
coral.backbone <- graph.data.frame(backbone2(coral.graph), 
                                   directed = F)
coral.backbone <- decompose.graph(coral.backbone)[[1]]
is_simple(coral.backbone)
coral.backbone <- simplify(coral.backbone, remove.multiple = T)
```


Veamos las propiedades de la red completa

#5. Network properties
##5.1 Basic properties

```{r}
V(coral.graph) #numero de nodos
E(coral.graph) #numero de enlaces
graph.density(coral.graph) #densidad de la red
count_components(coral.graph) #numero de componentes conectados
mean(igraph::degree(coral.graph)) #grado promedio
mean(strength(coral.graph))
```

##5.2 Backbone properties

```{r}
V(coral.backbone) #numero de nodos
E(coral.backbone) #numero de enlaces
graph.density(coral.backbone) #densidad de la red
count_components(coral.backbone) #numero de componentes
mean(degree(coral.backbone)) #grado promedio
mean(strength(coral.backbone))
```



##5.4 Distribucion de grado

```{r distribuciones de fuerza por atributo}
sd_att <- data.frame(a = glossary2$attribute , 
                     b = names(strength(coral.graph)), 
                     c =  strength(coral.graph),
                     row.names = seq(1,nrow(glossary2)
                                     )
                     )
```

```{r distribucion de fuerza de enfermedades}
plot_all <- ggplot(sd_att, aes(sd_att$c)) + 
  geom_histogram(binwidth = 40) + xlab("Vertex strength") +
  ylab("Frequency")
```


###5.4.1 Distribuciones particulares

```{r distribucion de fuerza de enfermedades}
sd_disease <- sd_att[sd_att$a=="disease",]
plot_disease <- ggplot(sd_disease, aes(sd_disease$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

```{r distribucion de fuerza de ecoregiones}
sd_ecoregion <- sd_att[sd_att$a=="ecoregion",]
plot_ecoregion <- ggplot(sd_ecoregion, aes(sd_ecoregion$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

```{r distribucion de fuerza de generos}
sd_genera <- sd_att[sd_att$a=="genera",]
plot_genera <- ggplot(sd_genera, aes(sd_genera$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```

```{r strenght distribution by topic}
sd_topic <- sd_att[sd_att$a=="topic",]
plot_topic <- ggplot(sd_topic, aes(sd_topic$c)) +
  geom_histogram(binwidth = 50) + xlab("Vertex strength") +
  ylab("Frequency")
```



```{r}
fig01top <- plot_grid(plot_all, plot_disease, labels = c("A", "B"), ncol = 2)
fig01mid <- plot_grid(plot_ecoregion, plot_genera, labels = c("C", "D"), ncol = 2)
fig01low <- plot_grid(plot_topic, labels = c("E"), ncol = 1)

svg("fig01.svg", width = 5)
plot_grid(fig01top, fig01mid, fig01low,  nrow = 3, align = "hv", 
          rel_widths = c(1, 1, 0.1)) #+ 
  #draw_label("DRAFT!", angle = 45, size = 80, alpha = .2)
dev.off()
```


#6. Top topics

```{r}
ecoregions <- tolower(coral.att$node[coral.att$attribute == "ecoregion"])
diseases <- tolower(coral.att$node[coral.att$attribute == "disease"])
genus <- tolower(coral.att$node[coral.att$attribute == "genera"])
topics <- tolower(coral.att$node[coral.att$attribute == "topic"])
```

```{r datos necesarios}
ls_ecoregions <- list()
ls_diseases   <- list()
ls_genus      <- list()
ls_topic      <- list()

for(i in 1:nrow(rawnet)) {
  ls_ecoregions[[i]] <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% ecoregions)]
  ls_diseases[[i]]   <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% diseases)]
  ls_genus[[i]]      <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% genus)]
  ls_topic[[i]]      <- rawnet[i, ][which(tolower(rawnet[i, ]) %in% topics)]
}

#number of papers by ecoregions
ecor_freq <- table(as.vector(t(do.call(cbind, ls_ecoregions)[1, ])))
#number of papers by diseases
enfe_freq <- table(as.vector(t(do.call(cbind, ls_diseases)[1, ])))
#number of papers by genera
gene_freq <- table(as.vector(t(do.call(cbind, ls_genus)[1, ])))
#number of papers by topic
topic_freq <- table(as.vector(t(do.call(cbind, ls_topic)[1, ])))
```


```{r top studied elements}
sum(tail(sort((ecor_freq*100/848)), 5))
sum(tail(sort((enfe_freq*100/848)), 6)[1:5])
sum(tail(sort((gene_freq*100/848)), 6)[1:5])
sum(tail(sort((topic_freq*100/848)), 6)[1:5])
```

```{r}
head(sort(ecor_freq, decreasing = T), 5)
head(sort(enfe_freq, decreasing = T), 6)
head(sort(gene_freq, decreasing = T), 6)
head(sort(topic_freq, decreasing = T), 5)
```


#7. Communities detection: edges communities


```{r link communities}
raw.link.com <- as.data.frame(cbind(get.edgelist(coral.backbone), 
                      E(coral.backbone)$weight))
colnames(raw.link.com) <- c("source","target","weight")
raw.link.com$weight <- as.numeric(as.character(raw.link.com$weight))
lc <- getLinkCommunities(raw.link.com)
```


```{r}
raw.link.com.id <- as.data.frame(cbind(get.edgelist(coral.backbone, names = F), 
                      E(coral.backbone)$weight))
colnames(raw.link.com.id) <- c("source","target","weight")
write.csv(raw.link.com, "coral.backbone.csv", row.names = F) #write edge list
write.csv(raw.link.com.id, "coral.backbone.id.csv", row.names = F) #write edge list
```


Now we analyze the relatedness between communities and determine possible metacommunities or adjustments in the cut value

```{r refining the communities}
cr <- getClusterRelatedness(lc) #closeness between communities
nested_comm <- getAllNestedComm(lc) #nested communities
#getNodesIn(lc, clusterids = c(18)) #membership of communities
```

```{r}
new_cut <- getMaxden(lc)
```

```{r recut}
lc2 <- (newLinkCommsAt(lc, cutat = new_cut[1])) # 11 comm, 0 nested
lc2$clustsizes
#write.csv(lc$nodeclusters, "node_membership.csv", row.names = F) #write node membership
#getNodesIn(lc2, clusterids = 1)
```


```{r community membership plot}
svg(filename="comm_membership.svg")
#win.metafile(filename="cairo", antialias="default")
comm.palette <- viridis_pal(option = "inferno")(lc2$numbers[3])
plotLinkCommMembers2(lc2, nodes = head(names(lc2$numclusters), 90),
                    pal = comm.palette,
                    total = T, fontsize = 6, sigmafont = 4, comfont = 5, sumfont = 4,
                    maxclusters = lc2$numbers[3], title = F, nspace = 5)
dev.off()
```



```{r external membership plot}
mmbr_matrix <- getCommunityMatrix(lc2, nodes = names(lc2$numclusters))
rownames(mmbr_matrix) <- names(lc2$numclusters)
write.csv(lc2$numclusters, file = "ext_nodes.csv")
write.csv(mmbr_matrix, file = "ext_members.csv")
```

now we can recheck the relatedness

```{r comm similarity}
#svg(filename="fig02C.svg")
cr2 <- getClusterRelatedness(lc2, hcmethod = "complete") #dissmililarity #between communities
cr2$height <- 100*cr2$height
#cr2$height <- 100-cr2$height
plot(cr2, main="", sub="", xlab="Communities", ylab="Similarity (%)", axes=F)
#axis(side = 2, at = seq(0,100,10), labels = seq(100,0,-10))
axis(side = 2, at = seq(0,100,10), labels = seq(100,0,-10))
#dev.off()
```

```{r jaccard similarity matrix}
cr2.matrix <- getClusterRelatedness(lc2, cluster = F) #Jaccard disimilarity
cr2.matrix <- 100*(1-cr2.matrix) #as % of Similarity
```


```{r simmilarity data}
comm.matrix <- getCommunityMatrix(lc2, nodes = head(names(lc2$numclusters), lc2$numbers[2]))
comm.matrix <- t(comm.matrix)
comm.jacc <- vegdist(comm.matrix,method = "jaccard", binary = T)
comm.jacc <- (1-comm.jacc)
comm.jacc <- as.matrix(comm.jacc)
comm.graph <- graph_from_adjacency_matrix(comm.jacc, mode = "lower", diag = F, weighted = T, 
                                          add.rownames = T)
E(comm.graph)$weight2 <- E(comm.graph)$weight*10
V(comm.graph)$sizes <- c(3, 3, 3, 5, 12, 4, 9, 22, 6, 37, 36)
V(comm.graph)$name <- c("C1", "C2", "C3", "C4", "C5", "C6",
                        "C7", "C8", "C9", "C10", "C11")
```


```{r figure02A-comm similarity graph}
#svg(filename="fig02A.svg")
comm_graph <- ggplot(data = ggnetwork(comm.graph, layout = "circle"),
       aes(x, y, xend = xend, yend = yend)) +
  geom_edges(aes(color = weight2, size = weight2*10)) +
  geom_nodes(aes(size = sizes**1.5), color = "black", fill = "black") +
  scale_color_gradient(low = "blue", high = "black") +
  geom_nodetext_repel(aes(label = vertex.names),
                      nudge_x = c(rep(0.05,6),rep(-0.05,5))) +
  scale_size_continuous(name = "Community\nsize",
                        breaks = 1.5*c(3, 9, 12, 22, 37),
                        labels = c("3", "9", "12", "22", "37"),
                        guide = guide_legend(override.aes = list(
                          linetype = c(rep(0, 5)))
                        )) +
  scale_color_continuous(name = "Shared nodes",
                         breaks = c(3/10, 7/10, 12/10, 17/10),
                         labels = c("3%", "7%", "12%", "17%"),
                         trans = "reverse") + theme_blank()
#dev.off()
```


```{r figure02B}
table_cnct <- as.data.frame(getCommunityConnectedness(lc2))
colnames(table_cnct) <- "Connectedness"
table_cnct$Community <- seq(1:lc2$numbers[3])
plot_cnct <- ggplot(table_cnct, aes(x = Community, y = Connectedness))+
  geom_bar(stat = "identity", col="black", fill = "steelblue") +
  scale_x_discrete(name = "Communities",
                   limits = seq(1:lc2$numbers[3]),
                   labels = c("C1", "C2", "C3", "C4", "C5", "C6",
                              "C7", "C8", "C9", "C10", "C11")) +
  ylab("Connectedness") + theme(axis.text=element_text(size = 10, angle = 90))
plot_cnct
```


```{r fig02 composed}
svg(filename = "fig02.svg", width = 8.5, height = 4)
plot_grid(comm_graph, plot_cnct, ncol = 2, labels = c("A", "B"), 
          rel_widths = c(1,0.7), label_size = 20)
dev.off()
```


#08. individual community plots

Each graph are assigned to a g.comm# where "#" is the number of the community

```{r, individual community plots}
for(i in 1:length(levels(lc2$edges$cluster))) {
  assign(paste("comm", levels(lc2$edges$cluster)[i], sep = ""), lc2$edges[lc2$edges[, 3] == i, ])
  assign(paste("g.", "comm", levels(lc2$edges$cluster)[i], sep = ""),
         ggnet2(graph.data.frame(get(paste("comm", levels(lc2$edges$cluster)[i], sep = ""))), label = T))
}
```


#09. Communities by trend

```{r small comms}
svg(filename = "fig03.svg")
plotLinkCommGraph(lc2, c(1, 2, 3, 4, 6), pal = comm.palette)
dev.off()
```

```{r mid sizes comms}
svg(filename = "fig04.svg")
plotLinkCommGraph(lc2, c(7, 9), pal = comm.palette)
dev.off()
```


```{r orbicella centered comm}
svg(filename = "fig05.svg")
plotLinkCommGraph(lc2, c(5))
dev.off()
```

```{r bbd centered comm}
svg(filename = "fig06.svg")
plotLinkCommGraph(lc2, c(8))
dev.off()
```

```{r acropora centered comm}
svg(filename = "fig07.svg")
plotLinkCommGraph(lc2, c(10))
dev.off()
```

```{r surveys comm}
svg(filename = "fig08.svg")
plotLinkCommGraph(lc2, c(11))
dev.off()
```

```{r whote network}
#svg(filename = "complete_network_test.svg", width = 15)
plotLinkCommGraph(lc2, vlabel = F, pal = comm.palette, scale.vertices = 0.25)
plotLinkCommGraph(lc2, vlabel = F, pal = comm.palette, scale.vertices = 0.25, 
                  layout = layout.circle)
plotLinkCommGraph(lc2, vlabel = F, pal = comm.palette, scale.vertices = 0.25, 
                  layout = layout.graphopt)
#dev.off()
```

External plots



```{r gephi-backbone}
sink("coral_backbone.gexf")
igraph.to.gexf(coral.backbone)
sink()
```


```{r gephi-main}
sink("coral_main.gexf")
igraph.to.gexf(coral.graph)
sink()
```



